---
title: "R Programming HW 5"
author: 'Rudecia Bernard'
format: 
  html:
    embed-resources: True
editor: visual
---

## Part A: Housing Data

Problem 1

```{r}

library(tidyverse)
houses <- read.csv("~/Downloads/R Homework and Datasets/Houses.csv")

colnames(houses)

model_1 <- lm(Total ~ SQFT, data = houses)

summary(model_1)

```

Problem 2:

The model predicts that home price increases by 135.74 for every increase in square feet.

Problem 3:

R-squared value is 0.02974, which asserts that about 3% of the variability in the data is due to square footage of the house.

Problem 4

```{r}

houses_p4 <- houses %>% 
  filter(Total != max(houses$Total))

model_2 <- lm(Total ~ SQFT, data = houses_p4)


summary(model_2)
```

Problem 5

Without the outlier, the R-squared increased dramatically. In the new model, approximately 67% of the variability in the data is accounted for by the square footage of the house.

Problem 6

```{r}


model_3 <- lm(Total ~ Acres + SQFT, data = houses)


summary(model_3)

```

Problem 7

The R-squared of model 3 is 0.9651, which is significantly higher than the version that just removed the outlier.

Problem 8

```{r}

t1 <- data.frame('Acres' = 0.5, 'SQFT' =2750)

predict(model_3, t1)

```

## Part B

Problem 1

```{r}

yt_a <- read.csv("~/Downloads/yt_comments_a.csv")
yt_b <- read.csv("~/Downloads/yt_comments_b.csv")


yt_a <- yt_a %>% 
  mutate(spam = as.character(if_else(spam == 1, 'yes', 'no')))
  


yt_a %>%
  count(spam)
```

Problem 2

```{r}

yt_a <- yt_a %>% 
  mutate(has_website = as.factor(str_detect(content, 'http')))
  
  
yt_a %>% 
  group_by(has_website, spam) %>% 
  summarize(count = n())

```

Problem 3

```{r}

library(rpart)

model_4 <- rpart(spam ~ has_website, data = yt_a)

model_4
```

Problem 4

```{r}

library(partykit)

plot.party(as.party(model_4))
```

Problem 5

```{r}

yt_a <- yt_a %>% 
  mutate(has_3_cons_nums = as.factor(str_detect(content, '\\d{3,}')),
         has_3_cons_punct = as.factor(str_detect(content, '[:punct:]{3,}')),
         has_sub = as.factor(str_detect(content, 'subscribe')))
                             
                             

glimpse(yt_a)


```

Problem 6

```{r}
model_5 <- rpart(spam ~ has_website + has_3_cons_nums + has_3_cons_punct + has_sub, data = yt_a)

 
model_5
 
plot.party(as.party(model_5))
```

Problem 7

```{r}

yt_b <- yt_b %>% 
   mutate(spam = as.character(if_else(spam == 1, 'yes', 'no')),
          has_3_cons_nums = as.factor(str_detect(content, '\\d{3,}')),
          has_3_cons_punct = as.factor(str_detect(content, '[:punct:]{3,}')),
          has_sub = as.factor(str_detect(content, 'subscribe')),
          has_website = as.factor(str_detect(content, 'http')))
                             
  

glimpse(yt_b)
```

Problem 8

```{r}

preds <- data.frame(predict(model_5, yt_b)) %>% 
  mutate(prediction = if_else(yes > no, 'yes', 'no')) %>% 
  select(prediction)


comp <- yt_b %>% 
  mutate(predictions = preds$prediction) %>% 
  group_by(predictions, spam) %>% 
  summarize(num = n())


comp

true_pos <- comp$num[2] + comp$num[3]


detected <- comp$num[3]

detected/true_pos


```

It seems like the model was very specific but not sensitive– there were no false positives, but there were quite a few spam comments that were not detected by the model.
